Article Title

### 2013-12-19 plos.tex
complete but untidy Discussion

\subsection*{Model selection}
Various estimation methods reveal radically different structures of functional connectivity? Which of the structures should be preferred? How does one determine which functional connectivity structure is best supported by the empirical data? Machine learning literature offers a rich toolbox of model selection criteria. Parametric criteria such as the \emph{Bayesian model selection criterion}, \emph{Akaike Information criterion}, and \emph{Bayes (or Schwartz) Information Criterion} are computationally efficient but rely on assumptions on generally unsupportable assumptions about the data generating process.  The most assumption-free yet more computationally intensive are procedures based on cross-validation \cite{Arlot:2010} and, when models have hyperparameters, nested cross-validation to selected the optimal values of the hyperparameters. It is commonly assumed that probabilistic models that demonstrate the highest average cross-validated log likelihood yield a better representation of the functional connectivity. For example, \cite{Ganmor:2011} show the fact that their proposed Reliable Network Model (a maximum-entropy model on the multivariate binary domain constrained by a sparse network of correlations of various orders) yielded higher cross-validated log-likelihoods than the full-pairwise Ising model. The authors use this fact to support the claim that the former model revealed a more anatomically relevant structure of functional connectivity than the pairwise model.

Under the loss function (Eq.~\ref{eq:loss}) employed in our study, the covariance matrix estimation methods are equivalent to fitting multivariate Gaussian models with various structure of dependencies between visible and inferred latent units and selecting between them on the basis of cross-validated log-likelihood. However, the ranking of models was preserved under other loss functions (Fig.~\ref{fig:S02}), indicating that model selection results were not strongly dependent on the assumption of gaussianity.

\subsection*{Big neural data}
Transformational breakthroughs in systems neuroscience are anticipated to follow the ongoing technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far \cite{Alivisatos:2013}. It is far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques.

Such probabilistic models of population activity feature large numbers of parameters, growing quadratically or faster with population size. Estimation of large number of parameters is sensitive to sampling noise. To solve this problem, the recording durations must increase at least proportionally to the number of parameters or the model must be constrained by regularization.

\subsection*{Future directions}

>>>>

\subsection*{Model selection}
Various estimation methods infer very different structures of functional connectivity from the same data. Which structure should be preferred? How does one determine which functional connectivity structure is best supported by the empirical data? The field of machine learning offers a rich toolbox of model selection criteria. Parametric criteria such as the \emph{Bayesian model selection criterion}, \emph{Akaike Information criterion} (AIC), and \emph{Bayes (or Schwartz) Information Criterion} (BIC) are computationally efficient but rely on generally unsupportable assumptions about the data generating process.  The most assumption-free yet more computationally intensive are procedures based on cross-validation \cite{Arlot:2010} and, when models have hyperparameters, nested cross-validation to select their values. It is commonly assumed that probabilistic models that demonstrate the highest average cross-validated log likelihood yield a better representation of the functional connectivity. For example, \cite{Ganmor:2011} demonstrated that their proposed Reliable Network Model (a maximum-entropy model on the multivariate binary domain constrained by a sparse heuristically selected network of correlations of several orders) yielded higher cross-validated log-likelihoods than the full-pairwise Ising model. The authors use this fact to support the claim that the former model revealed a more anatomically relevant structure of functional connectivity than the pairwise Ising model.

In simulation, we showed that, more constrained, simpler models often outperformed complex models in cross-validation studies even when their structure was differed from the true structure (Fig.~\ref{fig:02}\;row 5), only because these more constrained models were less sensitive to sampling noise. Therefore, by itself, better cross-validation performance of a model does imply closer correspondence of its structure to the functional structure of the data generating process. For example, a properly sparsified pairwise Ising model could outperform the Reliable Network model.  Unfortunately, optimal sparsification of Ising models is no computationally tractable procedure exists to infer the optimal sparse structure of dependencies in an Ising model.

In our study we selected between families of models based on their cross-validation performance. The winning model family was more flexible than the others and therefore more susceptible to noise.The fact that it still performed better indicated that the additional features were required. We conclude that under these circumstances, the selected model can be asserted to better represent the true structure of the data generating process, at least in comparison to the other three models.

In our study, we focused on estimating a specific parameter of the distribution (its covariance matrix) rather than fitting a specific parametric distribution.  Of course, under the loss function (Eq.~\ref{eq:loss}) we selected, the covariance matrix estimation methods are equivalent to fitting a multivariate Gaussian model and selecting the best structure on the basis of cross-validated log-likelihood. However, the ranking of models was preserved under other loss functions (Fig.~\ref{fig:Supp02}), indicating that model selection results were not strongly dependent on the assumption of gaussianity. We defend the practice of evaluating and optimizing probabilistic models with respect to their parameters that are more relevant for functional connectivity rather than maximizing their cross-validated log likelihoods, which are sensitive to all parameters. For example, model A may attain higher cross-validated log likelihood than model B due to its better ability to capture the tails of the distribution while model B better estimates the correlation matrix. We maintained that model B must be more relevant for studies of functional connectivity because correlations do reflect aspects of functional connectivity.

The correlation structures we considered also corresponded to structures of conditional dependencies in multivariate Gaussian models or, more generally, elliptical distributions. For example, zero partial correlation indicates conditional independence between a pair of variables under an elliptical distribution, but this correspondence weakens or breaks down with deviation from elliptical distributions \cite{Loh:2012}.  Other probabilistic models, when optimally sparsified for best cross-validated estimation of the covariance matrix, could suggest very different forms of low-dimensional representations (structures) of the covariance matrix. For example, spatiotemporal Ising models, marginalized to reproduce the correlation matrix at the same temporal scale, would suggest an alternative correlation structure from the same data. At this time we do not have computationally efficient methods for finding sparse structures of dependencies in Ising models optimized with respect to cross-validated estimation of their cross-correlation matrices. If we did, and the estimator thus constructed outperformed our $C_{\sf sparse+lowrank}$ estimator, we would be adopt this new estimator as one producing the best structure of the connectivity. Various forms of GLMs could also be evaluated with in the same way: sparsified for best estimation of the covariance matrix (not log-likelihood) and compared to the estimators studied in this study.  Again, if this new covariance estimator outperforms, $C_{\sf sparse+lowrank}$, we will be compelled to adapt it for estimating the functional connectivity.

\subsection*{Big neural data}
Transformational breakthroughs in systems neuroscience are anticipated to follow the ongoing technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far \cite{Alivisatos:2013}. It is far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques. Large datasets aggravate the numerical problems and the model selection caveats described in this paper. The number of parameters in models of population activity grows quadratically.  Models of joint population activity feature large numbers of parameters, that growing quadratically or faster with population size. Without ways to restrict such models, they quickly become ill conditioned (e.g. \cite{Roudi:2009}.  Statistical models of population activity will need to have both correct mathematical forms of interactions and optimal ways of restricting them so that the total number of estimated parameters remains commensurate with the amount of measurements.

Our study employs regularization, which restricts the models based on the data itself. Other models may use restrictions that are based on various heuristics in addition to regularization.  These methods have numerous caveats when it comes to inferring the optimal structure of functional connectivity from such methods. Improved estimation of covariances and the general framework of producing such estimators for specific neural populations can be applied to various circuits and serve as a statistical description of their functional organization.

\subsection*{Correspondence to anatomical organization}
If the ultimate goal of this line of inquiry to find descriptions of functional connectivity that are most informative about the anatomical and physiological organization of the circuit, then the models must be directly validated according to their correspondence to the anatomical organization. Little empirical evidence presently exists to show which models of connectivity, when fitted to the recordings of activity, reveal functional connectivity with best correspondence to anatomical connectivity.

In this study we find that the structure of partial correlation of the optimal estimator had stronger dependencies on the separation in depth, lateral position, and orientation preferences than sample correlations (Fig.~\ref{fig:07}\;A--C). The inferred positive and negative partial correlations too had distinct relationships to structure  (Fig.~\ref{fig:07}\;D--F).  Also the average partial correlations were more consistent between imaged sites than average sample correlations (Fig.~\ref{fig:06}\;A). Although these effects could conceivably results from trivial numerical artifacts, they are consistent with  improved representation of functional connectivity by the proposed estimator.

Additional studies with additional markers of circuit architecture such as cell types and synaptic connectivity could shed additional light on the ability of models to reveal meaningful functional connectivity.  For example, labeled interneurons could be studied for their distinct connectivity patterns contrasted with pyramidal cells.

\subsection*{Conclusions and future work}
We proposed an alternative statistically verified method for estimating the functional connectivity in neural circuits at temporal scales attainable by calcium imaging.
